---
title: "Course Project"
output:
  pdf_document: default
  html_document:
    keep_md: yes
  word_document: default
---

# The Problem Setting

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual transmissions

# The Data

First we load up the data, we've seen this data in a number of examples in lectures as well as quizzes. It has 11 variables with 32 observations of different car models. 

```{r, cache = TRUE, message = FALSE}
data(mtcars)
mtcars$vs <- as.factor(mtcars$vs); mtcars$am <- as.factor(mtcars$am) # should be categorical
```

Now let's make a plot the first question "Is an automatic or manual transmission better for MPG?":

```{r, message = FALSE}
library(ggplot2)
g <- ggplot(mtcars, aes(am, mpg))
g <- g + geom_boxplot(aes(fill=am))
g
```

Automatic is labeled as 1 and non automatic is labeled as 0. However there are a total of 10 predictors in this dataset, therefore there may be some interaction between them. Let's do an analysis of variance on all the predictors to see which play a bigger role in the determination of mpg:

```{r}
summary(aov(mpg ~ ., data = mtcars))
```

We can remove those models whose p-value is greater than 0.05. We do this in an automated way by doing:

```{r}
s <- summary(aov(mpg ~ ., data = mtcars))
# get the actual array info out of summary
s <- s[[1]]

# remove the row of residuals for now
idx <- s$`Pr(>F)`<0.05
rownames(s[idx, ])
```

So now let's update the model accordingly

```{r}
fit2 <- lm(mpg ~ cyl + disp + wt + drat + am, data = mtcars)
summary(fit2)
```

Again the same thing, refine the model

```{r}
fit3 <- lm(mpg ~ cyl + disp + wt + am, data = mtcars)
summary(fit3)
```

Now the last variable we remove is `disp`. Our final model is then

```{r}
final_fit <- lm(mpg ~ cyl + wt + am, data = mtcars)
summary(final_fit)
```

This says that we cannot reject the null hypothesis that the coefficient of am is 0. Let's do some diagnostic plots before making any conclusions, showing the Q-Q plot and residual plot among others:

```{r}
par(mfrow = c(2, 2))
plot(final_fit)
```

There is no real pattern in the residual plot which is very relieving. The Q-Q plot tells us that our assumption of normality is not violated. There is a rather large difference between MPG for automatic and manual transmissions. All in all, manual transmissions will get you more miles on the gallon.