---
title: 'T Confidence Intervals'
output:
  html_document:
    keep_md: yes
  word_document: default
---

# T Confidence Intervals

Recall previously we discussed creating confidence intervals using the central limit theorem. They are of the form

#### $Est \pm ZQ \cdot {SE}_{Est}$

where $Est$ is the estimate, $ZQ$ is the z distribution quantile, and ${SE}_{Est}$ is the standard error of the estimate. A t interval is applied when we have a small sample. They are of the form

#### $Est \pm TQ \cdot {SE}_{Est}$

where $TQ$ is the t distribution quantile. When given the choice of using the t vs z intervals, always go with t since you can always collect more data and the t quantile will approach that of the z quantile. 

The t interval works fairly well whenever the distribution of the data is roughly symmetric and mound shaped. Paired observations are often analyzed using the t interval by taking differences. When the data are skewed, you can work with the data on a log scale. For highly discrete data, other intervals are preferrable.

Let's take a look at Galton's sleep data, where he looked at sleep variations in subjects on different medications

```{r}
data(sleep)
sleep
```

We see there are two groups, with a subject id for each (1 - 10). 

## Results

```{r}
g1 <- sleep[sleep$group==1,]
g2 <- sleep[sleep$group==2,]

extra1 <- g1$extra
extra2 <- g2$extra

# compute the difference
difference = extra2 - extra1

# compute the mean and standard deviation of the difference
mn <- mean(difference)
s <- sd(difference)
n <- 10

# compute t confidence interval. qt gives quantile for t distribution
mn + c(-1, 1)*qt(.975, n-1)*s/sqrt(n)

# you can also get the t confidence intervals by running a t test on the difference
t.test(difference)

# or you can run a paired t test on both groups
t.test(extra1, extra2, paired = TRUE)
```

# Independent group t confidence intervals

Suppose we want to compare mean blood pressure between two groups in a randomized trial; those who received the treatment and those who received a placebo. We can't use the paired t test because the groups are independent and may have different sample sizes. The group confidence interval (t interval) in this case is given by

#### $\bar{Y} - \bar{X} \pm t_{n_x+n_y-2, 1- \alpha/2} \cdot S_p \cdot \left(\frac{1}{n_x} + \frac{1}{n_y} \right)^{1/2}$

where $t_{n_x+n_y-2, 1- \alpha/2}$ is the t quantile for the degrees of freedom of the combined samples ($n_x + n_y - 2$). Additionally, 

#### $S_p^2 = \frac{(n_x - 1)S_x^2 + (n_y - 1)S_y^2}{n_x+n_y-2}$ 

is called the ***pooled variance***. This interval assumes a constant variance across the two groups. If there is some doubt, assume a different variance per group (covered later).

Example: Chick weights

```{r}
library(datasets)
data("ChickWeight")
library(reshape2)

str(ChickWeight)
# define weight gain or loss
wideCW <- dcast(ChickWeight, Diet + Chick ~ Time, value.var = "weight")

# reshape based on time points in which the weight was measured and rename
names(wideCW)[-(1:2)] <- paste("time", names(wideCW)[-(1:2)], sep = "")
head(wideCW)

library(dplyr)
# create new variable which measures total weight gain/loss from beginning
# time point to end
wideCW <- mutate(wideCW, gain = time21 - time0)

# let's do a t interval
wideCW14 <- subset(wideCW, Diet %in% c(1, 4))

# bind the results with rbind
rbind(
t.test(gain ~ Diet, paired = F, var.equal = T, data = wideCW14)$conf, # equal variances
t.test(gain ~ Diet, paired = F, var.equal = F, data = wideCW14)$conf # unequal variances
)
```

In both cases the intervals are below zero, suggesting less weight gain on diet 1 than on diet 4. For the case when the variances are assumed to be ***unequal***, the t interval is given by 

#### $\bar{Y} - \bar{X} \pm t_{df} \cdot \left(\frac{s_x^2}{n_x} + \frac{s_y^2}{n_y} \right)^{1/2}$

If x and y are IID normal, the relevant statistic does not follow a t. Instead it can be *approximated* to be a t with a complicated degrees of freedom formula which is too long to typeset, nigga. But, as we saw above, all you need to do in `R` is specify `var.equal = F` and you're good.

When we have single observation, or paired observations which we take the diff, we use t intervals to create intervals highly robust to the underlying assumptions of the data. For count data, there's also the chi-squared tests and exact tests.