---
title: "Subsetting and Sorting"
output: 
  html_document:
    keep_md: true
---

Here's a quick review of how we generally subset and sort data frames. 

```{r}
set.seed(13435)
x <- data.frame("var1" = sample(1:5), "var2" = sample(6:10), 
                "var3" = sample(11:15))
x <- x[sample(1:5), ] # scramble the rows
x$var2[c(1, 3)] = NA # set values for rows 1 and 3 of var2 to be NA
x
```

To subset a column use 

```{r}
x[, 1]
```
or by column name:
```{r}
# or by colname
x[, "var1"]
```
specific rows pertaining to column of var2
```{r}
x[1:2, "var2"]
```
by logicals
```{r}
x[(x$var1 <= 3 & x$var2>1), ]
```
finding indices where a condition is true in the data frame:
```{r}
x[which(x$var2>8), ]
```

Here's how you order the rows of a data frame by a particular variable
```{r}
x[order(x$var1), ] # don't forget the comma
```

You can do the same thing with the `plyr` package and the `arrange` command:
```{r}
library(plyr)
arrange(x, var1)
```
or do it in descending order
```{r}
arrange(x, desc(var1))
```

Add a row by simply doing:
```{r sick}
x$var4 <- rnorm(5)

# or use cbind & make new df
y <- cbind(x, rnorm(5))
x
y
```
https://data.baltimorecity.gov/Community/Restaurants/k5ry-ef3g