# R Programming Assignment 3

These are my solutions to the third programming assignment of Coursera's R Programming Course. The instructions for this assignment can be found in the file "instructions.pdf" of this repo.

## 1. Plot the 30-day mortality rates for heart attack

```{r}
# name of files
fname_outcome <- "rprog-data-ProgAssignment3-data/outcome-of-care-measures.csv"
fname_hospital <- "rprog-data-ProgAssignment3-data/hospital-data.csv"

# loading outcomes
outcome <- read.csv(fname_outcome, colClasses="character")

# head() creates too much info, so let's use other tools
# print number of columns
ncol(outcome)

# print the name of the columns
names(outcome)
```

As you can see, there's a shitload of information in this file. Let's plot the 30 day death rates from heart attack (column 11). We need to coerce this as a numeric before doing so.

```{r}
outcome[, 11] <- as.numeric(outcome[, 11])
hist(outcome[, 11], col="red", breaks = 25, 
     main = "30-Day Mortality Rates From Heart Attack",
     xlab = "Mortality Rates",
     ylab = "Frequency")
rug(outcome[, 11])
abline(v = mean(outcome[, 11], na.rm=T), lwd=2, col="green")
```

## 2. Finding the best hospital in a state


