---
title: "Quiz 1"
output:
  html_document:
    keep_md: yes
  word_document: default
---

## Question 1 

```{r}
x <- c(0.18, -1.54, 0.42, 0.95)
w <- c(2, 1, 3, 1)
weighted.mean(x, w)
```

## Question 2 

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

lm(y ~ x - 1)$coefficients[[1]]
```

## Question 3 

```{r}
data("mtcars")

# wt is weight, in tons
lm(mpg ~ wt, mtcars)$coefficients[[2]]
```

## Question 4 

We have $Y$ and $X$. Given $\sigma(X) = 0.5 \cdot \sigma{Y}$, and $Corr(X, Y) = 0.5$. The value of the slope is given by 

$w_1 = Cor(Y, X) \cdot \frac{\sigma(y)}{\sigma(x)}$

So $w_1 = 0.5 * \frac{1}{0.5} = 1$

## Question 5 

$Cor(X, Y) = 0.4$ and both were normalized. When you normalize the variables to 0 mean and standard deviation of 1, the slope of the regression would be equivalent to the correlation, so you expect a score of $0.4*1.5 = 0.6$

## Question 6
```{r}
x <- c(8.58, 10.46, 9.01, 9.64, 8.86)
y <- (x - mean(x))/sd(x)
y[1]
```

# Question 7

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

lm(y ~ x)$coefficients[[1]]
```

## Question 8

Must be 0

## Question 9 

The mean
```{r}
mean(c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42))
```

## Question 10

The slope of a fit of $X$ to $Y$ is given by 

$\beta_1 = Cor(Y, X) \cdot \frac{\sigma(y)}{\sigma(x)}$

If you fit $Y$ to $X$ you get

$\gamma_1 = Cor(X, Y) \cdot \frac{\sigma(x)}{\sigma(y)}$

Given that $Cor(X, Y) = Cor(Y, X)$, then 

$\frac{\beta_1}{\gamma_1} = \frac{2\sigma(y)\sigma(y)}{2\sigma(x)\sigma(x)} = \frac{\sigma^2(y)}{\sigma^2(x)}$


